Political Polarization, Public Trust, and Healthcare Disparities: Unraveling COVID-19's Impact and Future Preparedness

### 2024-02-15 outputs/paper/paper.qmd
modified data section and introduction

Our analysis is grounded in comprehensive data sourced from the paper, `Why Did the Best Prepared Country in
the World Fare So Poorly during COVID?` [@citearticle], enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes. While leveraging the initial findings from the @citearticle, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities. We particularly focus on the impact of political polarization on adherence to health guidelines and the role of social vulnerabilities such as healthcare access, housing conditions, and ethnic demographics in shaping disparities in COVID-19 outcomes. One key finding that emerges from our study is the influence of political polarization, with Republican-leaning states exhibiting slightly higher clustering around elevated death rates, highlighting the importance of socio-political factors in shaping public health responses.

@@

As we navigate the complexities of the COVID-19 pandemic, it is imperative to not only understand the factors contributing to disparities but also to identify actionable strategies for mitigating these inequities and strengthening community resilience. In the subsequent sections of this paper, we first delve deeper into our methodology, data cleaning, and data measurement. Next, we present our findings in detail with graphs and tables. Finally, we discuss how these results came about and the implications of our research for public health policy and practice. By elucidating these critical relationships and expanding upon prior research, our study contributes to the broader discourse on pandemic preparedness and underscores the importance of tailored interventions in addressing underlying structural determinants of health disparities.

@@

The datasets utilized in the `original paper` [@citearticle] and this paper were mainly compiled from various sources such as the @data1, @data2, @data3, @data4, @data5, @data6, @data7, @data8, @data9, @data10, and @data11. Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and [@citedata2]. The political party support data we utilized from @citedata1 was compiled from @data12.

@@

This paper will replicate the data that was originally presented by @citearticle, as previously mentioned. We then transform the replication to our data analysis incorporating the COVID statistics and political party support data. We seek to use our initial replication and following own data analysis to provide a fresh part of the original paper, enhancing the with more data. Different from the original paper which spent a lot of time talking about how badly prepared the US was against other countries, We also specifically diverge our focus to the political and socioeconomic factors that impacted the US's COVID outcome.

@@

The transformation of real-world phenomena into quantifiable data points in our dataset is influenced by various factors, leading to inherent limitations in our  measurements for life expectancy rates [@citedefenition], and political affiliation.

Our measurement for COVID death rates is drawn from a accumulated dataset last updated until March 2023. This temporal constraint means we would not be aware of the fluctuations of death rates over time, particularly during the peak of the pandemic in 2020. The data's primarily reliance on the reported statistics from hospitals, clinics, testing centres, and laboratories may introduces bias as excludes uncounted deaths from COVID. Especially during the critical period of health facilities capacity constraints,many individuals died outside of medical facilities without being documented. Therefore, our measurements for COVID death rates are only accounted for the documented death from COVID.

@@

Similarly, our measurement of life expectancy suffer from the demographic limitations. Even though we categorize life expectancy by race, the absence of an age breakdown restricts us discern impacts across different age groups. Moreover, delays in reporting deaths undermine the timeliness and accuracy of our data, potentially skewing our understanding of population health trends.

@@

# Results

@@

# Discussion

@@

# Conclusion

>>>>

Our analysis is grounded in comprehensive data sourced from the paper, "Why Did the Best Prepared Country in the World Fare So Poorly during COVID?" [@citearticle], enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes. While leveraging the initial findings from the @citearticle, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities. We particularly focus on the impact of political polarization on adherence to health guidelines and the role of social vulnerabilities such as healthcare access, housing conditions, and ethnic demographics in shaping disparities in COVID-19 outcomes. One key finding that emerges from our study is the influence of political polarization, with Republican-leaning states exhibiting slightly higher clustering around elevated death rates, highlighting the importance of socio-political factors in shaping public health responses.

@@

As we navigate the complexities of the COVID-19 pandemic, it is imperative to not only understand the factors contributing to disparities but also to identify actionable strategies for mitigating these inequities and strengthening community resilience. In the subsequent sections of this paper, we first delve deeper into our methodology, data cleaning, and data measurement in @sec-data. Next, we present our findings in detail with graphs and tables in @sec-results. Finally, we discuss how these results came about and the implications of our research for public health policy and practice @sec-discussion and @sec-conclusion. By elucidating these critical relationships and expanding upon prior research, our study contributes to the broader discourse on pandemic preparedness and underscores the importance of tailored interventions in addressing underlying structural determinants of health disparities.

@@

The datasets utilized in the `original paper` [@citearticle] and this paper were mainly compiled from various sources such as the @data1, @data2, @data3, @data4, @data5, @data6, @data7, @data8, @data9, @data10, and @data11. Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and @citedata2. The political party support data we utilized from @citedata1 was compiled from @data12.

@@

This paper aims to replicate the data originally presented by @citearticle, as mentioned previously. Subsequently, we will integrate this replication into our own data analysis, incorporating COVID statistics and political party support data. Our objective is to leverage both the initial replication and our subsequent analysis to contribute a fresh perspective to the original paper, enriching it with additional data. In contrast to the original paper, which extensively discussed the inadequacies of US preparedness compared to other countries, we specifically diverge our focus to examine the political and socioeconomic factors that influenced the US's COVID outcomes.

@@

The transformation of real-world phenomena into quantifiable data points in our dataset is influenced by various factors, leading to inherent limitations in our  measurements for COVID death rates, life expectancy rates [@citedefenition], and political affiliation.

The determination of whether someone is deceased and whether they had COVID involves several steps and considerations. Medical professionals typically confirm a person's death through various means, including clinical examination, vital signs assessment, and sometimes laboratory tests. For COVID-related deaths, individuals are often tested for the presence of the virus through diagnostic tests such as PCR or antigen tests. Additionally, medical history and symptoms consistent with COVID-19 may also contribute to the determination of COVID-related fatalities. However, it's important to note that not all COVID-related deaths may be accurately identified or reported, leading to potential underestimation or misclassification in official statistics.

Incorporating this understanding into our analysis, our measurement for COVID death rates is drawn from a accumulated dataset last updated until March 2023. This temporal constraint means we would not be aware of the fluctuations of death rates over time, particularly during the peak of the pandemic in 2020. The data's primarily reliance on the reported statistics from hospitals, clinics, testing centres, and laboratories may introduces bias as excludes uncounted deaths from COVID. Especially during the critical period of health facilities capacity constraints,many individuals died outside of medical facilities without being documented. Therefore, our measurements for COVID death rates are only accounted for the documented death from COVID.

@@

Similarly, our measurement of life expectancy suffers from demographic limitations. Life expectancy, defined as the estimate of the average age that members of a particular population group will be when they die, is a crucial metric in understanding population health dynamics [@citedefenition]. Even though we categorize life expectancy by race, the absence of an age breakdown restricts our ability to discern impacts across different age groups. Moreover, delays in reporting deaths undermine the timeliness and accuracy of our data, potentially skewing our understanding of population health trends.

@@

# Results {#sec-results}

@@

# Discussion {#sec-discussion}

@@

# Conclusion {#sec-conclusion}
